Learning how to act: making good decisions with machine learning

### 2017-04-29 thesis/causality_tutorial.tex
bootcamp day2 update

We have developed a new drug for some illness and wish to determine how effective it is. We take a large group of patients and randomly assign half of them to a treatment group and the other half to a control group. The people in the treatment group get the drug, everyone else gets a placebo pill. The question we want to answer is does giving people the active drug improve their changes of recovery relative to giving them the placebo. We will use the variable $X$ (1 = drug, 0 = placebo) to represent the treatment each person receives and $Y$ (1 = recover, 0 = not recover) to describe the outcome .
\subsection{Structural Equation Models}
\label{sec:SEM}
Structural equation models (SEMs) describe a deterministic world, where underlying mechanisms determine the output of any process for a given input. The mechanism (but not the output) is assumed to be independent of what is fed into it. Linear structural equation models have a long history for causal estimation \cite {Wright1921,Haavelmo1943}. More recently, they have been formalized, generalized to the non-parametric setting and connected to developments in graphical models to provide a powerful causal framework \cite{Pearl2000}.

>>>>

We have developed a new drug for some illness and wish to determine how effective it is. We take a large group of patients and randomly assign half of them to a treatment group and the other half to a control group. The people in the treatment group get the drug, everyone else gets a placebo pill. The question we want to answer is does giving people the active drug improve their changes of recovery relative to giving them the placebo. We will use the variable $X$ (1 = drug, 0 = placebo) to represent the treatment each person receives and $Y$ (1 = recover, 0 = not recover) to describe the outcome.
\subsection{Structural Equation Models}
\label{sec:SEM}
Structural equation models (SEMs) describe a deterministic world, where underlying mechanisms determine the output of any process for a given input. The mechanism (but not the output) is assumed to be independent of what is fed into it. Linear structural equation models have a long history for causal estimation \cite {Wright1921,Haavelmo1943}. More recently, they have been formalized, generalized to the non-parametric setting and connected to developments in graphical models to provide a powerful causal framework \cite{Pearl2000}.

### 2017-04-29 thesis/thesis2.tex
bootcamp day2 update

@@

@@

>>>>

Key issue is that we can't observe the joint distribution over $Y_t$ and $Y_c$ even with randomized experiment.

@@

Note that we can bound this counterfactual distribution based on the variance of the observed interventional distributions.

Can these issues be resolved by considering personalized causal effects as random variables and avoiding counterfactuals?

Counterfactuals make a natural way to talk about causal effects in natural language. They could be seen as a short hand for projecting the world forward, holding everything else constant. An (impresise) way of expressing that we wish to consider the effect of only a single change.

What assumptions are required for the kind of counterfactual analysis like X would have been higher had B ...?

@@

Simpsons paradox
